Optimal design for on-farm strip trials --- systematic or randomised?

### 2024-02-08 mainV2_mark.tex
revised abstract and re-submit

\usepackage[url=false, isbn=false, eprint=false, backref=true, style= authoryear, backend=bibtex, maxcitenames=2, giveninits=true, maxbibnames=100]{biblatex}
\DeclareNameAlias{sortname}{family-given}
\addbibresource{BayesOFE.bib}

\title{Optimal design for on-farm strip trials --- systematic or randomised?}
%\author{Jerome}
\author[1]{Zhanglong Cao}
\author[1]{Jordan Brown}
\author[1,2]{Andrew Grose}
\author[1]{Mark Gibberd}
\author[1]{Julia Easton}
\author[1,2]{Suman Rakshit}

@@

\begin{abstract}
% There is no doubt on the importance of randomisation in agricultural experiments by agronomists and biometricians. Even when agronomists extend the experimentation from small trials to large on-farm trials, randomised designs predominate over systematic designs. However, the situation may change depending on the objective of the on-farm experiments (OFE). If the goal of OFE is obtaining a smooth map showing the optimal level of a controllable input across a grid made by rows and columns covering the whole field, a systematic design should be preferred over a randomised design in terms of robustness and reliability. With the novel geographically weighted regression (GWR) for OFE and simulation studies, we conclude that, for large OFE strip trials, the difference between randomised designs and systematic designs is not significant if a linear model of treatments is fitted or if the spatial variation is not taken into account. But for a quadratic model, systematic designs are superior to randomised designs. 
Randomisation is a crucial aspect of agricultural experiments for both agronomists and biometricians. While randomised designs are predominantly used for small trials and large on-farm experiments (OFE), the choice between randomised and systematic designs may depend on the objective of the OFE. Suppose the goal is to produce a smooth map of optimal input levels across a grid covering the entire field, in that case, a systematic design may be preferred for its robustness and reliability. Our simulation study using the Bayesian hierarchical model and geographically weighted regression (GWR) shows that, for large OFE strip trials, the difference between randomised and systematic designs is insignificant when fitting a linear model or ignoring spatial variation. However, for quadratic models, particularly with the presence of spatial variation, systematic designs are superior to randomised designs in terms of smaller true mean squared errors (MSE) of coefficients.
\end{abstract}

@@

We investigate the combination of the following factors: types of design with two levels: randomised and systematic; response relationship with two levels: linear and quadratic; correlation of coefficients with two levels: low and high; spatial variation with three levels: identity (no spatial trend), $\AR\otimes\AR$ and \Matern form; GWR bandwidth with three levels: 5, 9 and optimum given by AICc. The fixed bandwidth 5 in the simulation study covers all treatment levels (five nitrogen levels) in a systematic design, where the information is adequate in the inference of a quadratic curve. Similarly, the fixed bandwidth 9 covers all possible treatment levels in a randomised design. This is because if all treatments are randomly allocated in the strips, there is a chance that the information of one treatment level is missing if the bandwidth is less than 9.

@@

\begin{figure}[!htp]
	\begin{subfigure}[t]{0.45\textwidth}
		\centering
		\includegraphics[width=\linewidth]{Expt/Col_RandNitro.pdf}
		\caption{Treatments are randomly allocated into large strips in each replicate block.}\label{fig:RandNitrogen}
	\end{subfigure}
	\hspace{0.05\textwidth}
	\begin{subfigure}[t]{0.45\textwidth}
		\centering
		\includegraphics[width=\linewidth]{Expt/Col_SystNitro.pdf}
		\caption{Treatments are systematically allocated into large strips in each replicate block.}\label{fig:SystNitrogen}
	\end{subfigure}
	\caption{The nitrogen treatments with five levels (0, 35, 70, 105 and 140 kg/ha) randomly \eqref{fig:RandNitrogen} and systematically \eqref{fig:SystNitrogen} allocated into strips in each replicate block.}\label{fig:Nitrogen}
\end{figure}

@@

\begin{figure}[!htp]
	\begin{subfigure}[t]{0.45\textwidth}
		\centering
		\includegraphics[width=\linewidth]{Expt/LinlinesV2.pdf}
		\caption{Linear relationship between crop yield and nitrogen levels, where $y=65+0.05x$.}
	\end{subfigure}
	\hspace{0.05\textwidth}
	\begin{subfigure}[t]{0.45\textwidth}
		\centering
		\includegraphics[width=\linewidth]{Expt/QualinesV2.pdf}
		\caption{Quadratic relationship between crop yield and nitrogen levels, where $y=65+0.05x-0.003x^2$.}
	\end{subfigure}
	\caption{Noise-free linear and quadratic relationships between crop yield and nitrogen levels.}\label{fig:Lines}
\end{figure}

% Examples of the yield maps with true $\AR\otimes\AR$ spatially varying coefficients are in Figures \ref{fig:LinearYieldMap} and \ref{fig:QuadYieldMap}. 
% \begin{figure}[!htp]
% 	\centering
% 	\begin{subfigure}[t]{0.45\textwidth}
% 		\centering
% 		\includegraphics[width=\linewidth]{Expt/LinearRandYield01.pdf}
% 		\caption{Yield map of the linear response of the randomised design. }
% 	\end{subfigure}
% 	\begin{subfigure}[t]{0.45\textwidth}
% 		\centering
% 		\includegraphics[width=\linewidth]{Expt/LinearSystYield01.pdf}
% 		\caption{Yield map of the linear response of the systematic design. }
% 	\end{subfigure}
% 	\caption{Yield maps of linear response of randomly and systematically allocated nitrogen rates.}\label{fig:LinearYieldMap}
% \end{figure}

% \begin{figure}[!htp]
% 	\centering
% 	\begin{subfigure}[t]{0.45\textwidth}
% 		\centering
% 		\includegraphics[width=\linewidth]{Expt/QuadRandYield01.pdf}
% 		\caption{Yield map of the quadratic response of the randomised design. }
% 	\end{subfigure}
% 	\begin{subfigure}[t]{0.45\textwidth}
% 		\centering
% 		\includegraphics[width=\linewidth]{Expt/QuadSystYield01.pdf}
% 		\caption{Yield map of the quadratic response of the systematic design. }
% 	\end{subfigure}
% 	\caption{Yield maps of quadratic response of randomly and systematically allocated nitrogen rates.}\label{fig:QuadYieldMap}
% \end{figure}

@@

\begin{figure}[!htp]
	\centering
	\includegraphics[width=\linewidth]{Expt/Col_LinCombMSE_newpar.pdf}
	\caption{Boxplots of the logarithm of MSE for $\hat{\beta}_0$ and $\hat{\beta}_1$ in GWR models using different bandwidths for the simulated data with a linear response. The simulated data had different spatial covariance matrices (NS, AR1$\otimes$AR1 and \Matern) and a low correlation between the parameters ($\epsilon=1$).}\label{fig:LinBetaMSE}
\end{figure}

@@

\begin{figure}[!thp]
	\centering
	\includegraphics[width=\linewidth]{Expt/Col_QuaCombMSE_newpar.pdf}
	\caption{Boxplots of the logarithm of MSE for $\hat{\beta}_0$, $\hat{\beta}_1$ and $\hat{\beta}_2$ in GWR models using different bandwidths for the simulated data with a quadratic response. The simulated data had different spatial covariance matrices (NS, AR1$\otimes$AR1 and \Matern) and a low correlation amongst the parameters ($\epsilon=1$).} \label{fig:QuadBetaMSE}
\end{figure}

@@

\subsection{AICc Bandwidth selection}\label{Sec:bandselect}

@@

\begin{figure}[!htp]
	\begin{subfigure}[t]{0.45\textwidth}
		\centering
		\includegraphics[width=\linewidth]{Expt/BandwidthLinear.pdf}
		\caption{Histogram of optimal bandwidth for linear response.}
	\end{subfigure}
	\hspace{0.05\textwidth}
	\begin{subfigure}[t]{0.45\textwidth}
		\centering
		\includegraphics[width=\linewidth]{Expt/BandwidthQuadra.pdf}
		\caption{Histogram of optimal bandwidth for quadratic response.}
	\end{subfigure}
	\caption{Histogram of optimal bandwidth found by AICc for linear and quadratic response. }\label{fig:histband}
\end{figure}

@@

\section{Acknowledgements}

@@

@@

\begin{figure}[H]
	\centering
	\includegraphics[width=\linewidth]{Expt/Col_LinCombMSE_newpar_eta01.pdf}
	\caption{Boxplots of the logarithm of MSE for $\hat{\beta}_0$ and $\hat{\beta}_1$ in GWR models using different bandwidths for the simulated data with a linear response. The simulated data had different spatial covariance matrices (NS, AR1$\otimes$AR1 and \Matern) and a high correlation between the parameters ($\epsilon=0.1$).}\label{fig:LinBetaMSEeta01}
\end{figure}

\begin{figure}[H]
	\centering
	\includegraphics[width=\linewidth]{Expt/Col_QuaCombMSE_newpar_eta01.pdf}
	\caption{Boxplots of the logarithm of MSE for $\hat{\beta}_0$, $\hat{\beta}_1$ and $\hat{\beta}_2$ in GWR models using different bandwidths for the simulated data with a quadratic response. The simulated data had different spatial covariance matrices (NS, AR1$\otimes$AR1 and \Matern) and a high correlation amongst the parameters ($\epsilon=0.1$).} \label{fig:QuadBetaMSEeta01}
\end{figure}

>>>>

\usepackage[url=false, isbn=false, eprint=false, backref=true, style= authoryear, backend=bibtex, maxcitenames=2, giveninits=true, maxbibnames=100, uniquename=init]{biblatex}
%backend=bibtex
\DeclareNameAlias{sortname}{family-given}
\addbibresource{BayesOFE.bib}

\title{Optimal design for on-farm strip trials --- systematic or randomised?}
%\author{Jerome}
\author[1]{Zhanglong Cao}
\author[1]{Jordan Brown}
% \author[1,2]{Andrew Grose}
\author[1]{Mark Gibberd}
\author[1]{Julia Easton}
\author[1,2]{Suman Rakshit}

@@

% \begin{abstract}
% Randomisation is a crucial aspect for the estimation of unbiased treatment effects in agricultural experiments. While randomised designs are predominantly used for small trials and large on-farm experiments (OFE), the choice between randomised and systematic designs may depend on the objective of the OFE. Suppose the goal is to produce a smooth map of optimal input levels across a grid covering the entire field, in that case, a systematic design may be preferred for its robustness and reliability. Our simulation study using the Bayesian hierarchical model and geographically weighted regression (GWR) shows that, for large OFE strip trials, the difference between randomised and systematic designs is insignificant when fitting a linear model or ignoring spatial variation. However, for quadratic models, particularly with the presence of spatial variation, systematic designs are superior to randomised designs in terms of smaller true mean squared errors (MSE) of coefficients.
% \end{abstract}

% Field Crops Research has recently altered the format of the Abstract to make it more readable. I request that you revise the Abstract of your manuscript using the following headings and guidelines and resubmit your manuscript:

% A concise and factual abstract of no more than 400 words is required. It should report concisely the main findings of the research. To this end, the abstract is structured in five parts: 1. CONTEXT OR PROBLEM to present the background and issues; 2. OBJECTIVE OR RESEARCH QUESTION where the main hypothesis is presented; 3. METHODS to provide an overview of the materials and methods used; 4. RESULTS to present the main results using quantitative facts where possible; 5 CONCLUSIONS to present the main interpretation of the results; 6. IMPLICATIONS OR SIGNIFICANCE highlights the novelty of the results and their implications for field crops research and practice. An abstract is often presented separately from the article, so it must be able to stand alone. For this reason, references are not allowed. Also, non-standard or uncommon abbreviations should be avoided, but if essential they must be defined at their first mention in the Abstract. The headings must be used in the Abstract.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%% reviewers %%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Matthew J. Pringle,  matthew.pringle@qld.gov.au
%% Department of Environment and Science, GPO Box 2454, Brisbane, QLD 4001, Australia
%% OFE trial, experimental design for OFE

%% Prof. Dr. Hans-Peter Piepho, piepho@uni-hohenheim.de
%% Biostatistics (340c) [Director], Institute of Crop Science (340) [Managing director]
%% University of Hohenheim
%% OFE trial and model

%% Dr. Fiona Evans,  fiona@daa.com.au
%% https://www.daa.com.au/our-people/fiona-evans
%% GWR, simulation

%% Maria Lie Selle, maria.selle@ntnu.no
%% Department of Mathematical Sciences, Norwegian University of Science and Technology (NTNU), Trondheim, Norway
%% Bayesian model

%% Osval Antonio Montesinos-López, oamontes2@hotmail.com
%% Facultad de Telemática, Univ. de Colima, Colima, Colima, 28040 México
%% OFE trials experiments, Multivariate Bayesian Analysis of On-Farm Trials with Multiple-Trait and Multiple-Environment Data

\begin{abstract}

\zc{CONTEXT OR PROBLEM: Randomised designs are often preferred over systematic designs by agronomists and biometricians. However, for on-farm trials, the choice may depend on the objective of the experiments. If the purpose is to create a prescription map of a continuous input for each plot of a grid covering a large strip trial, a systematic design may be a better choice, but it attracts less discussion and attention. }

\zc{OBJECTIVE OR RESEARCH QUESTION: This study aims to evaluate the performance of systematic designs with geographically weighted regression (GWR) models in addressing spatial variation and estimating continuous treatment effects in large strip trials through numeric simulations.}

\zc{METHODS: A Bayesian hierarchical model with spatially correlated random parameters is utilised to generate simulated data for various scenarios of large strip on-farm trials. The study employs GWR models to analyse the simulated data for two assumptions: a linear response and a quadratic response to the treatment effects.}

\zc{RESULTS: With the assumption of linear response, the difference between a systematic design and a randomised design is not significant regardless of the presence of spatial variation. However, with the assumption of quadratic response, a systematic design is superior to a randomised design concerning the lower mean squared errors (MSE) by GWR. }

\zc{CONCLUSIONS: The findings highlight the superiority of systematic designs for producing smooth maps of optimal input levels and the impact of spatial variation on the performance of quadratic models in strip trials. Additionally, the study recommends fixed bandwidths based on experimental designs for GWR models. For a large strip trial to create a varying treatment map, a systemic design should be used as it has more flexibility in post-experiment statistical modelling.}

\zc{IMPLICATIONS OR SIGNIFICANCE: The findings offer practical recommendations for designing large strip trials. By drawing attention to the considerations, especially regarding the experiment's purpose, this research contributes valuable insights for improving the efficacy and planning of future large strip trials.}

\end{abstract}

@@

We investigate the combination of the following factors: types of design with two levels: randomised and systematic; response relationship with two levels: linear and quadratic; correlation of coefficients with two levels: low and high; spatial variation with three levels: identity (no spatial trend), $\AR\otimes\AR$ and \Matern form; GWR bandwidth with three levels: 5, 9 and optimum given by AICc. The fixed bandwidth 5 in the simulation study covers all treatment levels (five nitrogen levels) in a systematic design, where the information is adequate in the inference of a quadratic curve. Similarly, a fixed bandwidth of 9 is essential to cover all possible treatment levels in a randomised design. This necessity stems from situations where treatments are randomly distributed across strips. In particular, if identical treatments are positioned at the far left edge of the first replicate block and the far right edge of the second replicate block, a bandwidth smaller than 9 would lead to the GWR model capturing only the treatments between these boundaries, thereby missing the treatment level at the extremes.

@@

\begin{figure}[!htp]
	\begin{subfigure}[t]{0.45\textwidth}
		\centering
		\includegraphics[width=\linewidth]{Col_RandNitro.pdf}
		\caption{Treatments are randomly allocated into large strips in each replicate block.}\label{fig:RandNitrogen}
	\end{subfigure}
	\hspace{0.05\textwidth}
	\begin{subfigure}[t]{0.45\textwidth}
		\centering
		\includegraphics[width=\linewidth]{Col_SystNitro.pdf}
		\caption{Treatments are systematically allocated into large strips in each replicate block.}\label{fig:SystNitrogen}
	\end{subfigure}
	\caption{The nitrogen treatments with five levels (0, 35, 70, 105 and 140 kg/ha) randomly \eqref{fig:RandNitrogen} and systematically \eqref{fig:SystNitrogen} allocated into strips in each replicate block.}\label{fig:Nitrogen}
\end{figure}

@@

\begin{figure}[!htp]
	\begin{subfigure}[t]{0.45\textwidth}
		\centering
		\includegraphics[width=\linewidth]{LinlinesV2.pdf}
		\caption{Linear relationship between crop yield and nitrogen levels, where $y=65+0.05x$.}
	\end{subfigure}
	\hspace{0.05\textwidth}
	\begin{subfigure}[t]{0.45\textwidth}
		\centering
		\includegraphics[width=\linewidth]{QualinesV2.pdf}
		\caption{Quadratic relationship between crop yield and nitrogen levels, where $y=65+0.05x-0.003x^2$.}
	\end{subfigure}
	\caption{Noise-free linear and quadratic relationships between crop yield and nitrogen levels.}\label{fig:Lines}
\end{figure}

% Examples of the yield maps with true $\AR\otimes\AR$ spatially varying coefficients are in Figures \ref{fig:LinearYieldMap} and \ref{fig:QuadYieldMap}. 
% \begin{figure}[!htp]
% 	\centering
% 	\begin{subfigure}[t]{0.45\textwidth}
% 		\centering
% 		\includegraphics[width=\linewidth]{LinearRandYield01.pdf}
% 		\caption{Yield map of the linear response of the randomised design. }
% 	\end{subfigure}
% 	\begin{subfigure}[t]{0.45\textwidth}
% 		\centering
% 		\includegraphics[width=\linewidth]{LinearSystYield01.pdf}
% 		\caption{Yield map of the linear response of the systematic design. }
% 	\end{subfigure}
% 	\caption{Yield maps of linear response of randomly and systematically allocated nitrogen rates.}\label{fig:LinearYieldMap}
% \end{figure}

% \begin{figure}[!htp]
% 	\centering
% 	\begin{subfigure}[t]{0.45\textwidth}
% 		\centering
% 		\includegraphics[width=\linewidth]{QuadRandYield01.pdf}
% 		\caption{Yield map of the quadratic response of the randomised design. }
% 	\end{subfigure}
% 	\begin{subfigure}[t]{0.45\textwidth}
% 		\centering
% 		\includegraphics[width=\linewidth]{QuadSystYield01.pdf}
% 		\caption{Yield map of the quadratic response of the systematic design. }
% 	\end{subfigure}
% 	\caption{Yield maps of quadratic response of randomly and systematically allocated nitrogen rates.}\label{fig:QuadYieldMap}
% \end{figure}

@@

\begin{figure}[!htp]
	\centering
	\includegraphics[width=\linewidth]{Col_LinCombMSE_newpar.pdf}
	\caption{Boxplots of the logarithm of MSE for $\hat{\beta}_0$ and $\hat{\beta}_1$ in GWR models using different bandwidths for the simulated data with a linear response. The simulated data had different spatial covariance matrices (NS, AR1$\otimes$AR1 and \Matern) and a low correlation between the parameters ($\epsilon=1$).}\label{fig:LinBetaMSE}
\end{figure}

@@

\begin{figure}[!thp]
	\centering
	\includegraphics[width=\linewidth]{Col_QuaCombMSE_newpar.pdf}
	\caption{Boxplots of the logarithm of MSE for $\hat{\beta}_0$, $\hat{\beta}_1$ and $\hat{\beta}_2$ in GWR models using different bandwidths for the simulated data with a quadratic response. The simulated data had different spatial covariance matrices (NS, AR1$\otimes$AR1 and \Matern) and a low correlation amongst the parameters ($\epsilon=1$).} \label{fig:QuadBetaMSE}
\end{figure}

@@

\subsection{AICc bandwidth selection}\label{Sec:bandselect}

@@

\begin{figure}[!htp]
	\begin{subfigure}[t]{0.45\textwidth}
		\centering
		\includegraphics[width=\linewidth]{BandwidthLinear.pdf}
		\caption{Histogram of optimal bandwidth for linear response.}
	\end{subfigure}
	\hspace{0.05\textwidth}
	\begin{subfigure}[t]{0.45\textwidth}
		\centering
		\includegraphics[width=\linewidth]{BandwidthQuadra.pdf}
		\caption{Histogram of optimal bandwidth for quadratic response.}
	\end{subfigure}
	\caption{Histogram of optimal bandwidth found by AICc for linear and quadratic response. }\label{fig:histband}
\end{figure}

@@

\section*{Acknowledgements}

@@

\section*{CRediT authorship contribution statement}

\textbf{Zhanglong Cao}: Conceptualization; Formal analysis; Methodology; Software; Validation; Visualization; Roles/ Writing - original draft; and Writing - review \& editing. \textbf{Jordan Brown}: Investigation; Methodology; Visualization; Roles/Writing - original draft; and Writing - review \& editing. \textbf{Mark Gibberd}: Project administration; Roles/Writing - original draft; and Writing - review \& editing. \textbf{Julia Easton}: Roles/Writing - original draft; and Writing - review \& editing. \textbf{Suman Rakshit}: Conceptualization; Methodology; Supervision; Roles/Writing - original draft; and Writing - review \& editing.

@@

\begin{figure}[H]
	\centering
	\includegraphics[width=\linewidth]{Col_LinCombMSE_newpar_eta01.pdf}
	\caption{Boxplots of the logarithm of MSE for $\hat{\beta}_0$ and $\hat{\beta}_1$ in GWR models using different bandwidths for the simulated data with a linear response. The simulated data had different spatial covariance matrices (NS, AR1$\otimes$AR1 and \Matern) and a high correlation between the parameters ($\epsilon=0.1$).}\label{fig:LinBetaMSEeta01}
\end{figure}

\begin{figure}[H]
	\centering
	\includegraphics[width=\linewidth]{Col_QuaCombMSE_newpar_eta01.pdf}
	\caption{Boxplots of the logarithm of MSE for $\hat{\beta}_0$, $\hat{\beta}_1$ and $\hat{\beta}_2$ in GWR models using different bandwidths for the simulated data with a quadratic response. The simulated data had different spatial covariance matrices (NS, AR1$\otimes$AR1 and \Matern) and a high correlation amongst the parameters ($\epsilon=0.1$).} \label{fig:QuadBetaMSEeta01}
\end{figure}
